---
title: "Workings"
output:
  html_notebook:
    toc: yes
    toc_float: yes
  github_document:
    toc: yes
---

# Load libraries
```{r}
library(magrittr)
library(ggplot2)
```

# Merge datasets
Load datasets:
```{r cache = TRUE, results='hide'}
library(readr)

fun.df <- read_csv("./data/fundamentals.csv", 
    col_types = cols(`Period Ending` = col_date(format = "%Y-%m-%d"), 
        X1 = col_skip()))

quotes.df <- read_csv("./data/prices-split-adjusted.csv", 
    col_types = cols(date = col_date(format = "%Y-%m-%d")))

secs.df <- read_csv("D:/Code/R/Kaggle/nyse/data/securities.csv", 
    col_types = cols(`Date first added` = col_date(format = "%Y-%m-%d")))
```
Join `quotes.df` with `secs.df`:
```{r}
quotes.df %<>%
  dplyr::left_join(secs.df, by = c("symbol" = "Ticker symbol"))

write.csv(quotes.df, "./data/quotes_with_secs.csv", row.names = F)
```

## Merging fundamentals data with price data
Add period that the fundamentals data cover:

Note:

1. We assume data is available 90 days after a period end.

2. SEC actually has a range of [deadlines](https://en.wikipedia.org/wiki/Form_10-K#Filing_deadlines) to file the Form 10-K reports.
```{r}
fun.df %<>%
  dplyr::mutate(
  min.cover = `Period Ending` + 90,
  max.cover = c(`Period Ending`[-1], `Period Ending`[n()] + lubridate::dyears(1)) + 90
  )
```

Join the fundamentals data to the prices data:
```{r}
quotes.fun.df <- quotes.df %>%
  dplyr::left_join(fun.df, c("symbol" = "Ticker Symbol")) %>%
  dplyr::filter(date >= min.cover & date <= max.cover)

```

View a random sample to double check that the join makes sense:
```{r}
tgt.cols <- c("Period Ending", "min.cover", "max.cover")

quotes.fun.df %>%
  dplyr::sample_n(20) %>%
  dplyr::select(match(tgt.cols, names(.)),
                match(setdiff(names(.), tgt.cols), names(.)))
```

What are the quotes that are missing fundamentals data?
```{r}
x <- quotes.df %>%
  dplyr::select(date, symbol)

y <- quotes.fun.df %>%
  dplyr::select(date, symbol)

dplyr::setdiff(x, y) %>%
  dplyr::mutate(year = lubridate::year(date)) %>%
  dplyr::group_by(symbol, year) %>%
  dplyr::summarise(count = n())
```

How many stocks have fundamentals data?
```{r}
quotes.fun.df %>%
  dplyr::group_by(symbol) %>%
  dplyr::summarise(obs = n(), min.ob = min(date), max.ob = max(date)) %>%
  dplyr::arrange(obs) 
```

# Data Preprocessing
## Financial Ratios
Add some ratios computed off the fundamentals data:
```{r}
quotes.fun.df %<>%
  dplyr::mutate(`p/e ratio` = close/`Earnings Per Share`,
                `debt/equity ratio` = `Total Liabilities`/`Total Equity`)

write.csv(quotes.fun.df, "./data/quotes_with_fun.csv", row.names = F)
```

Sanity check the fundamentals data:
```{r}
quotes.fun.df %>%
  dplyr::select(-date:-`Period Ending`, -min.cover, -max.cover) %>%
  summary
```
Conclusion:

Data is not usable due to nonsensical values e.g. negative liabilities.

